An Airdrop That Preserves Recipient Privacy

### 2019-01-18 paper/rsa.tex
split paper into separate files per section

\documentclass[11pt]{article}
\usepackage[hidelinks]{hyperref}
\usepackage{amsthm,amsmath,amsfonts,amssymb,amstext,mathtools}
\usepackage{fullpage}
\usepackage{macros}

\begin{document}

\title{An Airdrop That Preserves Recipient Privacy}
\author{Dan Boneh \and Riad Wahby}

\maketitle

\begin{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

More to come.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Private airdrop: definitions}

A private airdrop (PAD) system is a triple of efficient algorithms:
\begin{itemize}
\item $\text{sendTokens}(\PK) \to (c, s)$:
Generates a token $c$ to be posted on the blockchain,
and a secret~$s$ given to the recipient via a private channel.
For example, an encryption of $s$ under $\PK$ can be posted
to the blockchain.   The posted encrypted $s$ is not associated with 
the token $c$.

\item $\text{sign}(\SK, (c, s), m) \to \text{sig}$:
Signs a message $m$ indicating that the token $c$ is spent.

\item $\text{verify}(c, m, \text{sig}) \to \{\text{yes},\text{no}\}:$
Verify that a signature $\text{sig}$ on token $c$ is valid.
\end{itemize}

TODO: Need to state security properties formally: 
(1) $c$ and $\text{sig}$ should reveal nothing about $\PK$,
(2) without $\SK$ we have existential unforgeability under a 
(one-time) chosen message attack.

\subsection{Zero-knowledge proofs in generic groups}

TODO: Need overview of zero-knowledge in generic groups.

\paragraph{The representation extraction lemma.}
Let $\adv$ be an algorithm that outputs 
a generic group element $u \in \GG$.
The following lemma shows that there is an extractor that
can extract from $\adv$ an integer representation of $u$
relative to a supplied set of group generators.
Moreover, this integer representation is unique.

\begin{lemma}[Unique representation extraction in generic groups]
\label{lem:unique}
Let $\adv_1, \adv_2$ be two randomized algorithms that interact
with group oracles for a generic group $\GG$ of unknown order.
For $i=1,2$, algorithm $\adv_i$ outputs $u_i \in \GG$.
Suppose that each algorithm makes at most $q$ type-1 queries
and let $g_1,\ldots,g_q \in \GG$ be the returned random group elements.

There is an extractor $\bdv$ that emulates 
the generic group oracles for $\adv_1$ and $\adv_2$
such that for $i=1,2$ when $\bdv$ interacts with $\adv_i$ 
the following holds with overwhelming probability:
if $\adv_i$ outputs $u_i \in \GG$ 
then the extractor $\bdv_i$ outputs a representation
$\alpha_{i,1},\ldots,\alpha_{i,q} \in \ZZ$
such that $u_i = g_1^{\alpha_{i,1}} \cdots g_q^{\alpha_{i,q}}$.
Moreover, if $u_1 = u_2$ then the two representations are the same,
namely $\alpha_{1,j} = \alpha_{2,j}$ for $j=1,\ldots,q$. 
\end{lemma}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A private airdrop to ECDSA and Schnorr public keys}

TODO: needs to be written.  Maybe put this first because it is so much
simpler?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A private airdrop to RSA public keys}

@@

To prove knowledge of the factorization of $n$ the prover
proves knowledge of a $w \in \range{N}$ such that $w^2 \equiv t \pmod{n}$
for some small public prime $t \in \ZZ$, namely $2 \leq t \leq \lambda$.
That is, we provide a ZKPK for the relation
\begin{equation} \label{eq:lang}
  {\cal R}_{g,h} \deq \left\{ \Bigl( (c, t) \in \GG \times \range{\lambda},\ \ 
                              (n, s, w, a) \in \range{N}^4 \Bigr)  \ \ 
        \text{s.t.}\ \ 
                \begin{split}
                        c   & = g^n \cdot h^s \ \ \text{in $\GG$}, \\
                        w^2 & = t + a \cdot n \ \ \text{in $\ZZ$}, \\
                        2 & \leq t \leq \lambda  \ \textrm{a prime}
                \end{split}  \right\}.
\end{equation}
Here $(c, t)$ is a statement and $(n, s, w, a)$ is a witness.
The integer relation $w^2 = t + a\cdot n$ proves that 
$w^2 \equiv t \pmod{n}$, as required.
We note that a ZKPK for~\eqref{eq:lang} proves that $t \in \ZZ$ is a quadratic
residue modulo the committed $n$.  Leaking this bit of information about $n$
does not interfere with the application to a secure airdrop.

@@

\begin{theorem}
Protocol $\prot{P}$ is a zero-knowledge protocol
for the relation ${\cal R}_{g,h}$ in \eqref{eq:lang}.
\end{theorem}

\begin{proof}
The simulator takes $(c, t)$ as input and outputs the simulated transcript
$(c_1, c_2, \Rv, \ell, \mathit{ch}, \Zv_q, \zv_\ell)$ as follows:
\begin{alg}
\> choose random $\hat{c}_1, \hat{c}_2 \rgets \GG,\ \ \ell \rgets \primes$, and
$\mathit{ch} \rgets \range{2^\lambda}$, \AlgSep
\> set $c_1 \gets \hat{c}_1^\ell$ and $c_2 \gets \hat{c}_2^\ell$, \AlgSep
\> choose random $\zv_\ell \rgets \range{\ell}^8$, \AlgSep
\lnum \label{ln:randv}
\> choose random $\hat{\rv} \rgets \range{2^{2\lambda}}^8$
and define $\hat{T} \deq \Big(\hat{c}_1^\mathit{ch}, \hat{c}_2^\mathit{ch}, 1, 1, \lfloor (t \cdot \mathit{ch})/\ell \rfloor\Big) \in \GG^4 \times \ZZ$, \AlgSep
\> set $\Zv_q \deq \hat{T} \cdot \phi(\lfloor \hat{\rv}/\ell \rfloor) \in \GG^4 \times \ZZ$, \AlgSep
\lnum \label{ln:computeR}
\> compute $\Rv \deq \Zv_q^\ell \cdot \phi(\zv_\ell) / T^\mathit{ch} \in \GG^4 \times \ZZ$, \AlgSep
\> output $(c_1, c_2, \Rv, \ell, \mathit{ch}, \Zv_q, \zv_\ell)$.
\end{alg}
We show the simulated transcript is distributed statistically close
to an accepting transcript of a real protocol execution.
Because $\ell$ does not divide $\abs{\GG}$ with overwhelming probability,
we know that $c_1$ and $c_2$ are uniform in $\GG$. 
Hence, observe that $c_1, c_2, \ell, \mathit{ch}$ are all distributed 
exactly as in a real transcript.  It remains to show that the same holds
for $\zv_\ell, \Zv_q$, and $\Rv$. 
In an interaction with a real prover 
the vector $\rv$ is chosen uniformly in $\range{2^{2\lambda}}^8$.
Consider one value $r \rgets \range{2^{2\lambda}}$.
Because $2^{2\lambda}$ is much larger than $\ell$
we know that the pair 
$\big(\lfloor r/\ell \rfloor,\ (r \bmod \ell) \big)$
is statistically close to uniform in $\range{2^{2\lambda}/\ell} \times \range{\ell}$.
It directly follows that $\zv_\ell$, which is masked by $(\rv \bmod \ell)$
in a real transcript, is statistically close to uniform 
in $\range{\ell}^8$, as in the simulation.

Next we argue that $\Zv_q \in \GG^4 \times \ZZ$ is simulated correctly. 
Write $\hat{\rv}$ generated on line~\eqref{ln:randv} in the simulation
as $\hat{\rv} = \big(r_\text{w}, r_\text{w2}, r_\text{s1}, r_\text{a}, r_\text{na}, r_\text{s1w}, r_\text{sa}, r_\text{s2} \big)$.
Then the simulated $\Zv_q$ is computed as:
\begin{equation} \label{eq:sim}
\Zv_q \deq \left(\begin{array}{l}
Z_1 \gets \hat{c}_1^\mathit{ch} \cdot g^{\lfloor r_\text{w}/\ell \rfloor} \cdot h^{\lfloor r_\text{s1}/\ell \rfloor} ,\quad
Z_2 \gets \hat{c}_2^\mathit{ch} \cdot g^{\lfloor r_\text{a}/\ell \rfloor} \cdot h^{\lfloor r_\text{s2}/\ell \rfloor} , \AlgSep
Z_3 \gets g^{\lfloor r_\text{w2}/\ell \rfloor} \cdot h^{\lfloor r_\text{s1w}/\ell \rfloor} / \hat{c}_1^{\lfloor r_\text{w}/\ell \rfloor},\quad
Z_4 \gets g^{\lfloor r_\text{na}/\ell \rfloor} \cdot h^{\lfloor r_\text{sa}/\ell \rfloor} / \hat{c}^{\lfloor r_\text{a}/\ell \rfloor}, \AlgSep
Z_5 \gets \lfloor (t \cdot \mathit{ch})/\ell \rfloor + \lfloor r_\text{w2}/\ell \rfloor - \lfloor r_\text{na}/\ell \rfloor 
\end{array}\right) \in \GG^4 \times \ZZ.  
\end{equation}
Let's start with $Z_5 \in \ZZ$.  
In a real transcript 
$Z_5 \deq \lfloor (\mathit{w2} \cdot \mathit{ch} + r_\text{w2})/\ell \rfloor -
          \lfloor (\mathit{na} \cdot \mathit{ch} + r_\text{na})/\ell \rfloor$.
Therefore, because $\mathit{w2} - \mathit{na} = t$, we have
that $Z_5$ in a real transcript satifies
\[    Z_5 = \lfloor (t \cdot \mathit{ch})/\ell \rfloor + \lfloor r_\text{w2} / \ell \rfloor - \lfloor r_\text{na} / \ell \rfloor - \Delta \]
where $\Delta \in \{0, 1\}$. 
In a real transcript $\lfloor r_\text{w2} / \ell \rfloor$ 
and $\lfloor r_\text{na} / \ell \rfloor$ 
are statistically close to fresh uniform values in $\range{2^{2\lambda}/\ell}$, 
as in the the simulated $Z_5$ in~\eqref{eq:sim}.
The only difference is that the real distribution is shifted by $\Delta$
where as the simulated distribution is not.  
Because $\Delta$ is small, $\Delta \in \{0,1\}$,
this difference only negligibly affects the statistical distance
between the real and simulated distributions. 
We thus conclude that the simulated $Z_5$ in~\eqref{eq:sim} 
is statistically close to its real distribution.

Essentially the same argument shows that $Z_1, Z_2, Z_3, Z_4$ are
also simulated correctly. 
For example, in a real transcript 
\begin{align*}
    Z_1 & \deq g^{\lfloor (\mathit{ch}\cdot w + r_\text{w}) / \ell \rfloor} \cdot 
             h^{\lfloor (\mathit{ch}\cdot \mathit{s1} + r_\text{s1}) / \ell \rfloor} = \\
       & =   g^{\lfloor (\mathit{ch}\cdot w) / \ell \rfloor} \cdot
             h^{\lfloor (\mathit{ch}\cdot s1) / \ell \rfloor} \cdot
             g^{\lfloor r_\text{w} / \ell \rfloor - \Delta_\text{w}} \cdot 
             h^{\lfloor r_\text{s1} / \ell \rfloor - \Delta_\text{s1}} =  \\
       & = (\tilde{c}_1)^\mathit{ch} \cdot
             g^{\lfloor r_\text{w} / \ell \rfloor} \cdot 
             h^{\lfloor r_\text{s1} / \ell \rfloor} \cdot 
             \left[ (g^{1/\ell})^{\delta_\text{w} - \ell \cdot \Delta_\text{w}} \cdot 
               (h^{1/\ell})^{\delta_\text{s1} - \ell \cdot \Delta_\text{s1}} \right].
\end{align*}
where $\Delta_\text{w}, \Delta_\text{s1} \in \{0,1\}$
and $\delta_\text{w} \deq ((\mathit{ch} \cdot w) \bmod \ell) \in \range{\ell}$
and $\delta_\text{s1} \deq ((\mathit{ch} \cdot \mathit{s1}) \bmod \ell) \in \range{\ell}$.
Hence, $Z_1$ in a real transcript is distributed as 
the simulated $Z_1$ in~\eqref{eq:sim}, where the only
difference is the quantity 
$\Delta \deq (g^{1/\ell})^{\delta_\text{w} - \ell \cdot \Delta_\text{w}} \cdot 
               (h^{1/\ell})^{\delta_\text{s1} - \ell \cdot \Delta_\text{s1}} \in \GG$. 
We claim that this $\Delta$ only negligibly affects the statistical distance
between the real and simulated distributions. 
{\bf TODO:} explain why.

@@

Because $a \neq 0$, the extractor can obtain $(n \bmod \ell)$
from the left equality in~\eqref{eq:third}, and thus extract $n \in \ZZ$ by
repeating this procedure over several $\ell$ and then using the chinese
remainder theorem.  
Similarly, it can extract $s \in \ZZ$ from
the right equality in~\eqref{eq:third}.   
It thus obtains a witness $(n, s, w, a)$ 
for the relation ${\cal R}_{g,h}$ in~\eqref{eq:lang}, as required.
\end{proof}

\subsection{The concrete airdrop system}

@@

\bibliographystyle{alpha}
\bibliography{paper}

\end{document}

>>>>

\section{A private airdrop to RSA public keys}\label{sec:rsa}

@@

To prove knowledge of the factorization of $n$ the prover
proves knowledge of a $w \in \range{N}$ such that $w^2 \equiv t \pmod{n}$
for some small public prime $t \in \ZZ$, namely $2 \leq t \leq \lambda$.
That is, we provide a ZKPK for the relation
\begin{equation} \label{eq:lang}
  \mathcal{R}_{g,h} \deq \left\{ \Bigl( (c, t) \in \GG \times \range{\lambda},\ \ 
                              (n, s, w, a) \in \range{N}^4 \Bigr)  \ \ 
        \text{s.t.}\ \ 
                \begin{aligned}
                        c   & = g^n \cdot h^s \ \ \text{in $\GG$}, \\
                        w^2 & = t + a \cdot n \ \ \text{in $\ZZ$}, \\
                        2 & \leq t \leq \lambda  \ \textrm{a prime}
                \end{aligned}  \right\}.
\end{equation}
Here $(c, t)$ is a statement and $(n, s, w, a)$ is a witness.
The integer relation $w^2 = t + a\cdot n$ proves that 
$w^2 \equiv t \pmod{n}$, as required.
We note that a ZKPK for~\eqref{eq:lang} proves that $t \in \ZZ$ is a quadratic
residue modulo the committed $n$.  Leaking this bit of information about $n$
does not interfere with the application to a secure
\RSW{``secure'' $\rightarrow$ ``private''?}
airdrop.

@@

\begin{theorem}
Protocol $\prot{P}$ is a zero-knowledge protocol
for the relation $\mathcal{R}_{g,h}$ in \eqref{eq:lang}.
\end{theorem}

\begin{proof}
The simulator takes $(c, t)$ as input and outputs the simulated transcript
$(c_1, c_2, \Rv, \ell, \mathit{ch}, \Zv_q, \zv_\ell)$ as follows:
\begin{alg}
\> choose random $c_1, c_2 \rgets \GG,\ \ \ell \rgets \primes$, and
\ \ $\mathit{ch} \rgets \range{2^\lambda}$, \AlgSep
\lnum \label{ln:randvell}
\> choose random $\zv_\ell \rgets \range{\ell}^8$, \AlgSep
\lnum \label{ln:randZ}
\> choose random $Z_1, Z_2, Z_3, Z_4 \rgets \GG$, \\
\lnum \label{ln:sim}
\> choose random $\hat{r}_\text{w2}, \hat{r}_\text{na} \rgets \range{2^{2\lambda}}$
and set $Z_5 \deq \lfloor (t \cdot \mathit{ch})/\ell \rfloor + \lfloor \hat{r}_\text{w2} / \ell \rfloor - \lfloor \hat{r}_\text{na} / \ell \rfloor \in \ZZ$, \\
\> set $\Zv_q \deq (Z_1, Z_2, Z_3, Z_4, Z_5)  \in \GG^4 \times \ZZ$, \AlgSep
\lnum \label{ln:computeR}
\> compute $\Rv \deq \Zv_q^\ell \cdot \phi(\zv_\ell) / T^\mathit{ch} \in \GG^4 \times \ZZ$, \AlgSep
\> output $(c_1, c_2, \Rv, \ell, \mathit{ch}, \Zv_q, \zv_\ell)$.
\end{alg}
We show the simulated transcript is distributed statistically close
to an accepting transcript of a real protocol execution.
In a real transcript we know that $c_1$ and $c_2$ are uniform in $\GG$. 
Hence, observe that the simulated $c_1, c_2, \ell, \mathit{ch}$ are all distributed 
exactly as in a real transcript.  It remains to show that the same holds
for $\zv_\ell, \Zv_q$, and $\Rv$. 
In an interaction with a real prover the quantities
$r_\mathit{s1}, r_\mathit{s1w}, r_\mathit{sa}, r_\mathit{s2}$
are chosen uniformly in $\range{N}$.
Consider one such value $r \rgets \range{N}$.
Because $N$ is much larger than $\abs{\GG} \cdot \ell$
we know that the pair
$\big(\lfloor r/\ell \rfloor \bmod \abs{\GG},\ (r \bmod \ell) \big)$
is statistically close to uniform in $\range{\,\abs{\GG}\,} \times \range{\ell}$.
Similarly,  in a real transcript
each of $r_w, r_\mathit{w2}, r_\mathit{na}, r_\mathit{sa}$
is chosen uniformly in $\range{2^{2 \lambda}}$ and therefore
each of these four values modulo $\ell$
is statistically close to uniform in $\range{\ell}$.
It follows that $\zv_\ell$, which is masked by $(\rv \bmod \ell)$
in a real transcript, is statistically close to uniform 
in $\range{\ell}^8$, as in line~\eqref{ln:randvell} in the simulation.

Next we argue that $\Zv_q \in \GG^4 \times \ZZ$ is simulated correctly.
Observe that each of the first four components of $\Zv_q$ is a group
element in $\GG$ that is randomized by
$h^{\lfloor r_\mathit{s1} / \ell \rfloor},\
 h^{\lfloor r_\mathit{s2} / \ell \rfloor},\
 h^{\lfloor r_\mathit{s1w} / \ell \rfloor},\
 h^{\lfloor r_\mathit{sa} / \ell \rfloor}$ respectively.
Because these four exponents are close to uniform in $\range{\,\abs{\GG}\,}$ and are
independent of everything else in the transcript, it follows that
in a real transcript $Z_1, Z_2, Z_3, Z_4$ are independent and close to uniform elements in $\GG$.  This is also how these elements are chosen 
on line~\eqref{ln:randZ} in the simulation.

It remains to argue that $Z_5 \in \ZZ$ is simulated correctly. 
In a real transcript this $Z_5$ is computed as 
$Z_5 \deq \lfloor (\mathit{w2} \cdot \mathit{ch} + r_\text{w2})/\ell \rfloor -
          \lfloor (\mathit{na} \cdot \mathit{ch} + r_\text{na})/\ell \rfloor$.
Because $\mathit{w2} - \mathit{na} = t$, we have
that $Z_5$ in a real transcript satisfies
\[    Z_5 = \lfloor (t \cdot \mathit{ch})/\ell \rfloor + \lfloor r_\text{w2} / \ell \rfloor - \lfloor r_\text{na} / \ell \rfloor + \Delta \]
where $\Delta \in \{0, \pm 1\}$. 
In a real transcript $\lfloor r_\text{w2} / \ell \rfloor$ 
and $\lfloor r_\text{na} / \ell \rfloor$ 
are statistically close to fresh uniform values in $\range{2^{2\lambda}/\ell}$, 
as in the simulated $Z_5$ on line~\eqref{ln:sim}.
The only difference is that the real distribution is shifted by $\Delta$
where as the simulated distribution is not.  
However this $\Delta$ only negligibly affects the statistical distance
between the real and simulated distributions. 
The reason is that in both distributions, 
every possible value in the range of $Z_5$, 
except possibly at the edges, 
is obtained from the same number of values 
for the pair $(r_\text{w2}, r_\text{na})$. 
%%  would be good to prove this as a general lemma.
We thus conclude that the simulated $Z_5$ in~\eqref{ln:sim} 
is statistically close to its real distribution.

@@

Because $a \neq 0$, the extractor can obtain $(n \bmod \ell)$
from the left equality in~\eqref{eq:third}, and thus extract $n \in \ZZ$ by
repeating this procedure over several $\ell$ and then using the chinese
remainder theorem.  
Similarly, it can extract $s \in \ZZ$ from
the right equality in~\eqref{eq:third}.   
It thus obtains a witness $(n, s, w, a)$ 
for the relation $\mathcal{R}_{g,h}$ in~\eqref{eq:lang}, as required.
\end{proof}

\subsection{The concrete airdrop system}\label{sec:rsa:airdrop}

@@
